Figures for Paper

### 2021-07-19 paper_figures.Rmd
Add some analysis
pfkp, rap1gap, etc

plot_list = list()
for(t in 1:length(lib_comps)){
  plt = clean_data_table %>%
     filter(tissue_clean == lib_comps[t]) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,y = log10(library_size),fill = disease_group2)) + 
    geom_boxplot(colour = "black", outlier.colour = NA) +
    geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
    facet_wrap(~ tissue_clean,scales = 'free_x') + 
    # ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
    ggpubr::stat_compare_means(label = "p.signif",
                               label.x.npc = "center",
                               method = "wilcox",
                               comparisons = list(
                                                  c("Control","ALS-TDP"),
                                                  c("ALS \n non-TDP","ALS-TDP"))) +
      ggpubr::stat_compare_means(label = "p.signif", 
                               label.x.npc = "center", 
                               method = "wilcox",
                               comparisons = list(
                                                  c("Control","FTLD-TDP"),
                                                  c("FTLD-TDP","FTLD \n non-TDP"))) + 
    labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
    theme(axis.text.x = element_text(size = 15)) + 
    ggpubr::theme_pubr() + 
    scale_fill_manual(values = force_colors) +
            theme(axis.ticks.x = element_blank(),
              axis.text.x = element_blank(),
              legend.position = "none")

@@

cowplot::plot_grid(plotlist = plot_list)
```

@@

for(t in 1:length(lib_comps)){

@@

@@

# Relationship between STMN2 and UNC13A CE PSI
## Scatter plot showing the correlation in non-log space in cortex UNC13A & STMN2 CE

@@

```{r als_ftd_genes_venn}
als_ftd_genomics = c("APP", "CHMP2B", "CSF1R", "DNAJC5","UNC13A",
                     "DNMT1", "EPM2A", "GRN", "ITM2B",
                     "MAPT", "NHLRC1", "NOTCH3", "PRNP", 
                     "PSEN1", "PSEN2", "TARDBP", "TBK1", 
                     "TYROBP", "UBQLN2", "VCP", "ATN1", 
                     "ATXN10", "ATXN1", "ATXN2", "C9orf72", 
                     "HTT", "JPH3", "NOP56", "TBP", "ALS2", 
                     "ANG", "DCTN1", "FIG4", "FUS", "HNRNPA1", 
                     "OPTN", "PFN1", "SETX", "SIGMAR1", "SLC52A2", 
                     "SLC52A3", "SOD1", "VAPB", "AR")

@@

ipsc_de[gene_name %in% als_ftd_risk_genes][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() 
```

>>>>

plot_list = list()
for(t in 1:length(lib_comps)){
  plt = clean_data_table %>%
     filter(tissue_clean == lib_comps[t]) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,y = log10(library_size),fill = disease_group2)) + 
    geom_boxplot(colour = "black", outlier.colour = NA) +
    geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
    facet_wrap(~ tissue_clean,scales = 'free_x') + 
    # ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
    ggpubr::stat_compare_means(label = "p.signif",
                               label.x.npc = "center",
                               method = "wilcox",
                               comparisons = list(
                                                  c("Control","ALS-TDP"),
                                                  c("ALS \n non-TDP","ALS-TDP"))) +
      ggpubr::stat_compare_means(label = "p.signif", 
                               label.x.npc = "center", 
                               method = "wilcox",
                               comparisons = list(
                                                  c("Control","FTLD-TDP"),
                                                  c("FTLD-TDP","FTLD \n non-TDP"))) + 
    labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
    theme(axis.text.x = element_text(size = 15)) + 
    ggpubr::theme_pubr() + 
    scale_fill_manual(values = force_colors) +
            theme(axis.ticks.x = element_blank(),
              axis.text.x = element_blank(),
              legend.position = "none") + 
    ylab(element_blank())

@@

cowplot::plot_grid(plotlist = plot_list,nrow = 3)
```

@@

cell_types = fread(paste0(here::here(),"/data/All_1917_samples_Darmanis_dtangle_deconv.tsv"))
for(t in 1:length(lib_comps)){

@@

# Detection rate as a linear model
```{r}

@@

# Relationship between STMN2 and UNC13A CE PSI
## Scatter plot showing the correlation in non-log space UNC13A & STMN2 CE

```{r}
clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
  # filter(STMN2_cryptic_leaf >= 2) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
  geom_point() +
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
```
## Scatter plot showing the correlation in non-log space RAP1GAP1 & STMN2 CE

@@

```{r als_ftd_genes_venn}
als_ftd_genomics = c("ALS2", "ANG", "APP", "AR", "ATN1", "ATXN1", "ATXN10", "ATXN2", 
"ATXN3", "C21orf2", "C9orf72", "CHMP2B", "CSF1R", "DCTN1", "DNAJC5", 
"DNMT1", "EPM2A", "FIG4", "FUS", "GRN", "HNRNPA1", "HTT", "ITM2B", 
"JPH3", "KIF5A", "LOC101927815", "MAPT", "NHLRC1", "NOP56", "NOTCH3", 
"OPTN", "PFN1", "PMP22", "PRNP", "PSEN1", "PSEN2", "SCFD1", "SETX", 
"SIGMAR1", "SLC52A2", "SLC52A3", "SOD1", "TARDBP", "TBK1", "TBP", 
"TNIP1", "TYROBP", "UBQLN2", "UNC13A", "VAPB", "VCP")

@@

ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% clipr::write_clip()
```

## Linear model on UNC13A CE PSI with covariates
```{r the_silly_linear}
# library size 
#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
full_cov = clean_data_table %>%
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    filter(sequencing_platform != "") %>% 
    left_join(cell_types %>% 
                  dplyr::select(cell,deconv,sample) %>% 
                  unique() %>% 
                  pivot_wider(id_cols = "sample",names_from = "cell",values_from = "deconv"), by = 'sample') %>%
    filter(!is.na(disease_group2)) %>% 
    unique() %>% 
    filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
    dplyr::select(c("unc13a_cryptic_leaf_psi",
                    "sample",  "tissue_clean",  
                    "rin",  "sex", 
                    "age",  "mutations", 
                    "sequencing_platform", 
                    "number_g_alleles", 
                    "stmn_2_cryptic_psi_leaf", "oligodendrocytes", 
                    "astrocytes", "neurons", "endothelial", "microglia")) %>% 
    column_to_rownames('sample')

fit <- lm(unc13a_cryptic_leaf_psi ~ ., data = full_cov)

fit %>% broom::tidy() %>% fwrite(paste0(here::here(),"/covariate_unc13a_ce.csv"))
full_cov %>% 
    group_by(tissue_clean) %>% 
    nest() %>% 
    mutate(lm_obj = map(data, ~lm(unc13a_cryptic_leaf_psi ~., data = .x))) %>% 
    mutate(broomed = map(lm_obj, broom::tidy)) %>% 
    dplyr::select(tissue_clean,broomed) %>% unnest() %>% filter(p.value < 0.05) %>% View()

nested %>% 
    pluck("lm_obj", 1) %>% broom::tidy()
unc13a_data_fisher = unc13a_sequencing_platform %>% 
    filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
    group_by(disease_group2,tissue_clean) %>% 
    nest() %>% 
    mutate(fisher_pvalue = map_dbl(data, quick_fisher))

# Correlation between STMN2 CE PSI and RAP1GAP1/PFKP

```{r}
patient_others = fread(file.path(here::here(),"data","nygc_pfkp_unc13b_rap1gap.bed"))

patient_others[,sample := gsub(".SJ.out","",V4)]
clean_data_table = clean_data_table %>% 
    left_join(patient_others %>% 
                  pivot_wider(id_cols = "sample",names_from = "V7",values_from = "V5",values_fill = 0))

clean_data_table[, rap1gap_cryptic_psi := (RAP1GAP_noveldonor + RAP1GAP_novelacceptor)/ (RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated)]
clean_data_table[, pfkp_cryptic_psi := (PFKP_novelacceptor + PFKP_noveldonor)/ (PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated)]

pfkp_plot = clean_data_table %>%
    filter(grepl("Cortex",tissue_clean)) %>%
    filter(disease_tissue == T) %>%
    filter((PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated) >=30) %>% 
    filter(junction_reads_stmn2 >= 30) %>%
    filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(pfkp_cryptic_psi > 0) %>%
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = pfkp_cryptic_psi)) +
    geom_point() +
    stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
    geom_smooth(method = lm, se = FALSE,color = 'black') +
    xlab("STMN2 Cryptic PSI") +
    ylab("PFKP Cryptic PSI") +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 18)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1L))

rap1gap_plot = clean_data_table %>%
    filter(disease_tissue == T) %>%
    filter(grepl("Cortex",tissue_clean)) %>%
    filter((RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated) >=30) %>% 
    filter(junction_reads_stmn2 >= 30) %>%
    filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(rap1gap_cryptic_psi > 0) %>%
    # ggplot(aes(x = rap1gap_cryptic_psi, y = unc13a_cryptic_leaf_psi)) +
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = rap1gap_cryptic_psi)) +
    geom_point() +
    stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
    geom_smooth(method = lm, se = FALSE,color = 'black') +
    xlab("STMN2 Cryptic PSI") +
    ylab("RAP1GAP Cryptic PSI") +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 18)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1L))

cowplot::plot_grid(rap1gap_plot,pfkp_plot)
```

```{r reproduce_4a_in_rap1gap}
detected_table_for_graph_rap1 = clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(rap1gap_cryptic_psi > 0) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter((RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated) >=30) %>% 
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,rap1gap_cryptic_psi) %>%
  mutate(log2_rap1gap = log2((rap1gap_cryptic_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()

comparisons = combn(unique(detected_table_for_graph_rap1$detection_name),2,simplify = F)

detected_table_for_graph_rap1 %>%
  ggplot(aes(y = log2_rap1gap, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means() +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio RAP1GAP CE PSI / STMN2 CE PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
